---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data
```{r load, echo=TRUE}
library(lubridate)
setwd("~/RepData_PeerAssessment1")
activitydata <- read.csv("activity.csv")
activitydata$date <- ymd(activitydata$date)
```


## What is mean total number of steps taken per day?
```{r totalsteps, echo = TRUE}
t1 <- aggregate(steps ~ date, activitydata, sum)
hist(t1$steps, main = "Histogram -- Total Steps Taken Each Day", col = "blue")
m1 <- mean(t1$steps)
m2 <- median(t1$steps)
```
#### _The mean number of steps taken each day `r m1`._
#### _The median number of steps taken each day `r m2`._

## What is the average daily activity pattern?
```{r avgdailypattern, echo = TRUE}
plot(t1$date, t1$steps, type = "l", xlab = "Date", ylab = "Steps", main = "Time-Series of Daily Steps", col = "red")

maxstep <- t1[which(t1$steps==max(t1$steps)),1]
```
#### _The maximum average daily steps was recorded on `r maxstep`._


## Imputing missing values
```{r imputing, echo=TRUE}
ad1 <- activitydata
missingDataSum <- sum(is.na(ad1$steps))
t2 <- aggregate(steps ~ interval, ad1, median)

ad1$steps[which(is.na(ad1$steps))] <- t2$steps[which(t2$interval ==                                         ad1$interval[which(is.na(ad1$steps))])]

t3 <- aggregate(steps ~ date, ad1, sum)
hist(t3$steps, main = "Histogram -- Total Steps Taken Each Day (Imputed)", col = "blue")

m3 <- mean(t3$steps)
m4 <- median(t3$steps)

diff1 <- mean(t3$steps) - mean(t1$steps)
diff2 <- median(t3$steps) - median(t1$steps) 
```
#### _There are `r missingDataSum` missing data points in the original dataset._
#### _Strategy used to impute missing values --> Take median of 5 minute interval._
#### _The mean number of steps taken each day using imputed strategy `r m3`._
#### _The median number of steps taken each day using imputed strategy `r m4`._
#### _The new strategy resulted in change in mean by `r diff1` and median by `r diff2`._


## Are there differences in activity patterns between weekdays and weekends?
```{r weekend, echo=TRUE}
library(chron)
library(lattice)
ad1$weekend <- as.factor(is.weekend(ad1$date))
levels(ad1$weekend) <- c("Weekday", "Weekend")

## Scratch
weekdayData <- subset(ad1, ad1$weekend == "Weekday")
weekendData <- subset(ad1, ad1$weekend == "Weekend")

wd1 <- aggregate(steps ~ interval, weekdayData, mean)
we1 <- aggregate(steps ~ interval, weekendData, mean)

we1$weekend <- TRUE
wd1$weekend <- FALSE

test <- rbind(wd1, we1)
test$weekend <- as.factor(test$weekend)
levels(test$weekend) <- c("Weekday", "Weekend")

xyplot(steps ~ interval | weekend, data = test, layout = c(1,2), type = "l")
```

#### Looks like people start their days earlier on weekdays.  The activity peaks early, then trails off.  While weekend activty starts later, but activity level stays more or less the same thru the day.